\chapter{Installing Epsilon}
\label{chp:InstallingEpsilon}

There are two options for installing Epsilon. Through the Eclipse Update Manager or by checking out the source code from the Eclipse SVN and (if you'd like to play with experimental stuff) EpsilonLabs SVN servers. This chapter provides detailed instructions for both installation options.

\section{Prerequisites}

You first need to install the following software before you can install Epsilon.

\begin{enumerate}
	\item Java 1.5 (or later)
	\item Eclipse 3.3 (or later)
	\item EMF 2.3 (or later) 
	\item GMF 2.0 (or later) (optional)
	\item Emfatic (optional) (recommended)
\end{enumerate}

The Java runtime can be downloaded from \url{http://java.sun.com}. The easiest way to download Eclipse, EMF and GMF together is the Eclipse Ganymede Modeling distribution \url{http://www.eclipse.org/downloads/packages/eclipse-modeling-tools-includes-incubating-components/ganymeder}.

Emfatic is a textual notation for specifying ECore metamodels. It is highly recommended that you install it as all the exemplar metamodels in this book will be progressively specified using it (as opposed to using UML-like diagrams). Emfatic can be installed using the Eclipse Update Manager (a detailed description follows) using the update site : \url{http://epsilonlabs.svn.sourceforge.net/svnroot/epsilonlabs/org.eclipse.emf.emfatic.updatesite/}

\section{Installing using the Eclipse Update Manager/P2}

To install Epsilon using the Eclipse Update Manager please follow the instructions below:

\begin{enumerate}
	\item Go to Help->Software Updates
 	\item In the dialog that pops up press Add Site…
	\item In the dialog that pops up enter  \url{http://download.eclipse.org/modeling/epsilon/updates} in the URL field.
	\subitem In addition to the main update site above, there is also an interim site which hosts the current (but potentially unstable) version of Epsilon: \url{http://download.eclipse.org/modeling/epsilon/interim}
	\item Press OK
	\item In the next screen select Epsilon and press Next
	\subitem If you don't have GMF installed, expand the Epsilon node and deselect the Epsilon GMF Support feature
	\item Review the Feature Licence and if you agree select the I accept the terms in the licence agreement and press Next
	\item Press Finish
	\item Eclipse prompts for a restart. Press Yes
\end{enumerate}

\section{Checking out the Source Code from CVS/SVN}

\begin{enumerate}
	\item Switch to the SVN Repository Exploring perspective (from the Window$\rightarrow$Open Perspective$\rightarrow$Other menu)
	\item Click the New Repository Location button on the SVN Repositories view and enter the following configuration details
	\subitem URL : http://dev.eclipse.org/svnroot/modeling/org.eclipse.gmt.epsilon/
	\subitem User : anonymous
	\subitem Password : leave empty
	\item Open the trunk node
	\item  Select the sub-folders of plugins and features, right-click and check them out.
\end{enumerate}

\section{Which Option Should I Use?}

By checking out the source code, you will be able to pick up the latest updates and bug-fixes as soon as they are committed to the SVN. The downside is that you then have to launch a new Eclipse workbench where Epsilon will be available. On the other hand, it is easier to install using the Update Manager and you don't need to launch a new workbench every time, but you'll have to wait for the next binary release to get the latest features.